---
title: "Signing up for an openai API key"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this tutorial, we will guide you through the process of obtaining an API key from openai, which is necessary for using openai's large language models (LLMs) such as `gpt-4o` or `gpt-5` with the `quallmer` package. We will also cover how to configure your R environment to use the openai API key securely.

# Precautions 

**Closed LLMs such as OpenAI's GPT-4o are not free to use and require a subscription. They also come with ethical concerns and risks, especially when it comes to data privacy and security. Therefore, always be aware of the data you use and the potential consequences of your analysis and make sure to enable the necessary safeguards to protect privacy and security. In addition, be aware that the license to use [OpenAI models](https://openai.com/policies/row-terms-of-use/) comes along with adhering to specific regulations to avoid misuse. Therefore, always be aware of the data you use and the potential consequences of your analysis and make sure to enable the necessary safeguards to protect privacy and security.**


# Setting up an API key for openai models

The openai API provides access to various so called closed models (fee-based, not open-source). There are other services available (for example, Claude or Google Gemini, etc.) which require slightly different set-ups. You can find more information on pricing of openai [here](https://platform.openai.com/pricing>). 

### To use the openai API, please follow these steps:

1.  Go to the openai playground: [https://platform.openai.com/playground>](https://platform.openai.com/playground)
2.  Click on `Sign up` in the top right corner
3.  Fill in your details and confirm `Sign up`
4.  Click now on the `Settings` icon in the top right corner
5.  Go to `Billing` and **provide your billing information** (otherwise it won't work!)
6.  Once your billing information is complete, you can create a `new project` (top left corner, click on `Default Project`)
7.  Click on `Dashboard` in the top right corner
8.  Click on `API keys` in the left side panel at the bottom
9.  Click on `Create new secret key`
10. **Copy** the key before you close the window

Once you copied the key, you should **save it somewhere safe and accessible.** For security reasons, you won’t be able to view it again through your openai account. If you lose it, you will need to regenerate it. Keep your API key safe and do not share it with others. If you suspect that your key has been compromised, you can regenerate it in the dashboard. **Be aware that you will be charged for any usage of the API via your key.**

### Configuring your openai API key in RStudio

To interact with the openai API, it's required to have a valid `OPENAI_API_KEY` environment variable in R. You can establish this environment variable globally by including it in the so-called .Renviron file. This approach ensures that the environment variable persists across all your R sessions as the Shiny app runs in the background. Here is a set of commands to open the .Renviron file for modification:

```{r}
#| eval: false
require(usethis)
edit_r_environ()
```

Add the following line to .Renviron, replacing "APIKEY" with your actual API key: OPENAI_API_KEY="APIKEY". You need to restart your R session for the changes to take effect. You can do this by clicking on the `Session` menu in RStudio and selecting `Restart R`.

**Caution:** If you're using version control systems like GitHub or GitLab, remember to include .Renviron in your .gitignore file to prevent exposing your API key! To maintain the privacy of your data when using gptstudio, do not highlight, include in a prompt, or otherwise upload any sensitive data, code, or text that should remain confidential.

Now you are ready to use openai models with the `quallmer` package! For example, you can [test the setup by running the example on sentiment analysis.](https://quallmer.github.io/quallmer/articles/pkgdown/examples/sentiment.html)
