---
title: "Getting started with quallmer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with quallmer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Our tutorials provide a brief introduction to the `quallmer` package, which is designed to facilitate the use of large language models (LLMs) for qualitative research tasks. The package relies on the `ellmer` package for LLM interactions, providing a seamless interface for users to work with different LLM providers. For more information on the `ellmer` package and supported LLM interactions, please refer to its documentation [here](https://ellmer.tidyverse.org/index.html).

## Basic usage

To get started with `quallmer`, you first need to install the package from GitHub. 

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("SeraphineM/quallmer")
```

Then, you can load the package and begin using its functions. 
```{r setup}
library(quallmer)
```

## Overview of tutorials

Before using large language models, most users will need to sign up for an API key from an LLM provider, such as openai, or download an open-source model like those from Ollama. The `quallmer` package supports multiple LLM providers through the `ellmer` package, allowing users to choose the one that best fits their needs.

Using `annotate()`, users can generate structured, interpretable outputs powered by large language models (LLMs). The package includes a library of [predefined tasks for common qualitative coding needs](https://seraphinem.github.io/quallmer/articles/pkgdown/examples/overview.html), such as sentiment analysis, thematic coding, and stance detection. It also allows users to [create their own custom annotation tasks tailored to their specific research questions and data types](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/customtask.html) using `define_task()`. Our tutorials will guide you through the following topics:

1. [**Signing up for an OpenAI API key**](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/openai.html): This tutorial will guide you through the process of obtaining an API key from OpenAI, which is necessary for using OpenAI's LLMs with the `quallmer` package.

2. [**Working with an open-source Ollama model**](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/ollama.html): This tutorial will demonstrate how to use the `quallmer` package with an open-source Ollama model for qualitative coding tasks.

3. [**Creating custom tasks**](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/customtask.html): This tutorial will walk you through the process of defining your own custom annotation tasks using the `define_task()` function, allowing you to tailor the LLM's output to your specific research needs.

More tutorials will be added soon. 
