---
title: "Example: Ideology detection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `qlm_code()` function with the built-in `task_ideology()` codebook allows you to perform ideological scaling (0-10) on texts regarding a specified ideological dimension. In this example, we will demonstrate how to use the `task_ideology()` codebook for ideology detection on a sample corpus of innaugural speeches from U.S. presidents. We will use the dimension "inclusive–exclusive" as an example. To refine the codebook, we will also provide a definition of the dimension (optional).

### Loading packages and data

```{r getting-started}
# We will use the quanteda package 
# for loading a sample corpus of innaugural speeches
# If you have not yet installed the quanteda package, you can do so by:
# install.packages("quanteda")
library(quanteda)
library(quallmer)

# For educational purposes, 
# we will use a subset of the inaugural speeches corpus
# The three most recent speeches in the corpus
data_corpus_inaugural <- quanteda::data_corpus_inaugural[57:60]

```
### Using `qlm_code()` for ideological scaling of texts
```{r ideology}
# Define ideological dimension
dimension <- "inclusive–exclusive"
# Provide definition for the dimension
definition <- "Inclusive language emphasizes equal rights, diversity, pluralism,
and protection of minorities, whereas exclusive language emphasizes exclusion
of groups, national homogeneity, and restricting rights."
# Apply predefined ideology codebook with task_ideology() in the qlm_code() function
result <- qlm_code(data_corpus_inaugural, codebook = task_ideology(dimension, definition),
                   model = "openai/gpt-4o",
                   temperature = 0)
```
```{r display_results, echo = FALSE}
# results table
library(kableExtra)
result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3)
```
### Adjusting the ideology scaling task
You can customize the ideological scaling task by defining your own codebook with `qlm_codebook()` (for a more detailed explanation, [see our "Defining custom tasks" tutorial](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/customtask.html)).
For example, you might like to change the scale from 0-10 to -5 to +5.
```{r custom_ideology}
custom_ideology <- qlm_codebook(
    name = "Ideological scaling",
    instructions = paste0(
      "Task:",
      "- Read each short text carefully.",
      "- Place the text on a -5 to +5 scale for the following ideological dimension: ",
      dimension,
      definition
    ),
    schema = ellmer::type_object(
      score       = ellmer::type_integer("Ideological position on the specified dimension (-5 to +5, where -5 = first pole, +5 = second pole)"),
      explanation = ellmer::type_string("Brief justification for the assigned score, referring to specific elements in the text")
    ),
    role = "You are an expert political scientist performing ideological text scaling.",
    input_type = "text"
  )
# Apply the custom codebook
custom_result <- qlm_code(data_corpus_inaugural, codebook = custom_ideology,
                          model = "openai/gpt-4o",
                          temperature = 0)
```
```{r display_results_custom, echo = FALSE}
# custom results table
library(kableExtra)
custom_result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3)
```

In this example, we demonstrated how to use the `task_ideology()` codebook for scaling texts regarding their ideological position on a specified dimension. We also showed how to customize the codebook using the `qlm_codebook()` function for more tailored annotation needs, e.g., changing the scale from 0-10 to -5 to +5. Now you can apply these techniques to your own text data for ideological analysis!
