---
title: "Example: Sentiment analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `annotate()` function with a predefined `sentiment()` task allows you to rate the sentiment of texts using an LLM. The predefined `sentiment()` object structures the response with a numeric sentiment score from -1 (very negative) to 1 (very positive) and a brief explanation. 

### Loading packages and data

```{r getting-started}
library(quallmer)

#Example texts
texts <- c(
"This is wonderful!",
"I really dislike this approach.",
"The results are somewhat disappointing.",
"Absolutely fantastic work!"
)
```

### Using `annotate()` for predefined sentiment analysis of texts

```{r sentiment}
# Apply predefined sentiment task with task_sentiment() in the annotate() function
result <- annotate(texts, task = task_sentiment(), 
                   chat_fn = chat_openai, model = "gpt-4o",
                   api_args = list(temperature = 0, seed = 42))
```


```{r display_results2, echo = FALSE}
# sentiment analysis results table
library(kableExtra)
result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3) 
```

### Adjusting the sentiment task

You can customize the sentiment analysis task by defining your own task with `task()` (for a more detailed explanation, [see our "Defining custom tasks" tutorial](https://seraphinem.github.io/quallmer/articles/pkgdown/tutorials/customtask.html)).

For example, you might want to include an additional field for confidence level.
```{r custom_sentiment}
custom_sentiment <- task(
  name = "Custom sentiment analysis",
  system_prompt = "You are an expert annotator. Rate the sentiment of each text from -1 (very negative) to 1 (very positive), briefly explain why, and provide a confidence level from 0 to 1.",
  type_def = ellmer::type_object(
    score = ellmer::type_number("Sentiment score between -1 (very negative) and 1 (very positive)"),
    explanation = ellmer::type_string("Brief explanation of the rating"),
    confidence = ellmer::type_number("Confidence level from 0 to 1")
  ),
  input_type = "text"
)
# Apply the custom sentiment task
custom_result <- annotate(texts, task = custom_sentiment, 
                          chat_fn = chat_openai, model = "gpt-4o",
                          api_args = list(temperature = 0, seed = 42))
```
```{r display_results_custom, echo = FALSE}
# sentiment analysis results table
library(kableExtra)
custom_result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3) 
```
Or, you might want to change the scoring scale to a 5-point Likert scale.
```{r likert_sentiment}
likert_sentiment <- task(
  name = "Likert scale sentiment analysis",
  system_prompt = "You are an expert annotator. Rate the sentiment of each text on a scale from 1 (very negative) to 5 (very positive) and briefly explain why.",
  type_def = ellmer::type_object(
    score = ellmer::type_number("Sentiment score between 1 (very negative) and 5 (very positive)"),
    explanation = ellmer::type_string("Brief explanation of the rating")
  ),
  input_type = "text"
)
# Apply the Likert scale sentiment task
likert_result <- annotate(texts, task = likert_sentiment, 
                          chat_fn = chat_openai, model = "gpt-4o",
                          api_args = list(temperature = 0, seed = 42))
```
```{r display_results_likert, echo = FALSE}
# sentiment analysis results table
library(kableExtra)
likert_result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3) 
```
In this way, you can easily adapt the sentiment analysis task to fit your specific research needs!
